Sampling Clint and Funbound.plasma

### 2025-08-01 data-raw/ClearanceProject-0-Distinguishability.Rmd

```{r library_loading}
# June 6, 2025 8AM: Using httk-dev/HTTK-293 branch
# June 6, 2025 3PM: Using httk-dev/dev branch
source("data-raw/RC_comparison_bkgd.R")
library(dplyr)
library(tidyselect)
library(ggplot2)
library(patchwork)
library(parallel)
```

@@

```{r set_parameters}
THIS_CHEMICAL = "DTXSID3024289" # A chemical to test
# THIS_CHEMICAL = "DTXSID50873382" # Warfarin
DOSE = 25
SPECIES = "Rat"
ROUTE = "oral"
TIMES = c(seq(0, 23.75, by = 0.25), seq(24, 72, by = 6),
          seq(96, 120, by = 24))/24
SAMPLES = 1000
VARPARS = list(Clint = 0.1)
CENSORPARS = list(Funbound.plasma = list(cv = 0.1, lod = 1E-5))
HUMAN_CLINT_FUP = TRUE
SPECIES_TEXT <- ifelse(SPECIES != "Human" && HUMAN_CLINT_FUP,
                       paste0(SPECIES, "_parHuman"),
                       SPECIES)
INVITROUV = NULL
INVITROUV_TEXT = NULL
if (!is.null(INVITROUV)) {
  INVITROUV_TEXT = "_usedINVITRO"
}
```

```{r get_chems}
chems_available <- get_cheminfo(info = "dtxsid", species = SPECIES,
                                model = "gas_pbtk") |> suppressWarnings()
```

@@

@@

@@

```{r}
lombardo <- readxl::read_xlsx(
  paste0(Sys.getenv("FIG_DIR"),
         "Lombardo2018-Supplemental_82966_revised_corrected.xlsx"),
  skip = 8) |> janitor::clean_names()

table(
  lombardo$cas_number %in% suppressWarnings(get_cheminfo("CAS", model = "gas_pbtk"))
)

@@

temp_data3 <- left_join(temp_data2, temp_lst, relationship = "many-to-one")
saveRDS(temp_data3,
        paste0("data-raw/", SPECIES_TEXT, "_HTTK_preds_CvTdata.rds")
)
```

```{r bin-compare}
# Get the bins
bindf_rat <- readRDS("data-raw/Rat_oral_simulation_bins.rds") |>
  subset(select = c("DTXSID", "Bin"))
bindf_rat_parhuman <- readRDS("data-raw/Rat_parHuman_oral_simulation_bins.rds") |>
  subset(select = c("DTXSID", "Bin"))

@@

temp_data3 |>
  filter(Sol_R > 0, Sol_N > 0) |>
  mutate(FE_R = log10(Sol_R / Conc), FE_N = log10(Sol_N / Conc)) |>
  group_by(DTXSID, Species, Media, Route, Dose) |>
  summarize(AFE_R = 10^(mean(FE_R)), AFE_N = 10^(mean(FE_N)), N_Obs = n()) |>
  ungroup() |>
  left_join(bindf_rat_parhuman) |>
  mutate(Bin = ifelse(Bin == "Distinguished", "Distinguishable", Bin)) |>
  arrange(DTXSID, Species, Media, Route, Dose) |>
  ggplot(aes(x = AFE_N, y = AFE_R)) +
  geom_point(aes(fill = log10(Dose)), shape = 21, color = "black", size = 1.2, alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0) +
  geom_vline(xintercept = 1, linetype = "22", color = "grey20") +
  geom_hline(yintercept = 1, linetype = "22", color = "grey20") +
  facet_wrap(~Bin) +
  scale_fill_viridis_c() +
  labs(x = "Nonrestrictive Solution", y = "Restrictive Solution", 
       title = afe_formula) +
  scale_x_log10(labels = scales::label_log()) + scale_y_log10(labels = scales::label_log()) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "bottom",
        legend.title.position = "top")

@@

>>>>

```{r library_loading}
# June 6, 2025 8AM: Using httk-dev/HTTK-293 branch
# June 6, 2025 3PM: Using httk-dev/dev branch
source("data-raw/RC_comparison_bkgd.R")
library(dplyr)
library(tidyselect)
library(ggplot2)
library(patchwork)
library(parallel)
library(mirai)
library(ctxR)
```

@@

```{r set_parameters}
THIS_CHEMICAL = "DTXSID3024289" # A chemical to test
DOSE = 25
SPECIES = "Rat"
ROUTE = "oral"
TIMES = c(seq(0, 23.75, by = 0.25), seq(24, 72, by = 6),
          seq(96, 120, by = 24))/24
SAMPLES = 1000
VARPARS = list(Clint = 0.1)
CENSORPARS = list(Funbound.plasma = list(cv = 0.1, lod = 1E-5))
HUMAN_CLINT_FUP = TRUE
SPECIES_TEXT <- ifelse(SPECIES != "Human" && HUMAN_CLINT_FUP,
                       paste0(SPECIES, "_parHuman"),
                       SPECIES)
INVITROUV = NULL

INVITROUV_TEXT = NULL
if (!is.null(INVITROUV)) {
  INVITROUV_TEXT = "_usedINVITRO"
}
```

```{r get_chems}
chems_available <- get_cheminfo(info = "dtxsid", species = SPECIES,
                                model = "gas_pbtk", suppress.messages = TRUE)
```

@@

#### Testing

@@

#### Set-up assessment functions

@@

```{r lombardo_sipes}

load(file = "data-raw/CYP_Sipes2017.Rdata")

lombardo <- readxl::read_xlsx(
  paste0(Sys.getenv("FIG_DIR"),
         "Lombardo2018-Supplemental_82966_revised_corrected.xlsx"),
  skip = 8) |> 
  janitor::clean_names() |>
  left_join(chem.physical_and_invitro.data[c("CAS", "DTXSID")],
            by = join_by(cas_number == CAS))

table(
  lombardo$DTXSID %in% suppressWarnings(get_cheminfo("DTXSID", model = "gas_pbtk"))
)

@@

temp_data3 <- left_join(temp_data2, temp_lst, relationship = "many-to-one")
saveRDS(temp_data3,
        paste0("data-raw/RCdata/", SPECIES_TEXT, "_HTTK_preds_CvTdata.rds")
)
```

```{r bin-compare}
# Get the bins
bindf_rat <- readRDS("data-raw/RCdata/Rat_oral_simulation_bins.rds") |>
  subset(select = c("DTXSID", "Bin"))
bindf_rat_parhuman <- readRDS("data-raw/RCdata/Rat_parHuman_oral_simulation_bins.rds") |>
  subset(select = c("DTXSID", "Bin"))

@@

temp_data3 |>
  filter(Sol_R > 0, Sol_N > 0) |>
  mutate(FE_R = log10(Sol_R / Conc), FE_N = log10(Sol_N / Conc)) |>
  group_by(DTXSID, Species, Media, Route, Dose) |>
  summarize(AFE_R = 10^(mean(FE_R)), AFE_N = 10^(mean(FE_N)), N_Obs = n()) |>
  ungroup() |>
  left_join(bindf_rat_parhuman) |>
  mutate(Bin = ifelse(Bin == "Distinguished", "Distinguishable", Bin)) |>
  arrange(DTXSID, Species, Media, Route, Dose) |>
  ggplot(aes(x = AFE_N, y = AFE_R)) +
  geom_point(aes(fill = log10(Dose)), shape = 21, color = "black", size = 1.2, alpha = 0.6) +
  geom_abline(slope = 1, intercept = 0) +
  geom_vline(xintercept = 1, linetype = "22", color = "grey20") +
  geom_hline(yintercept = 1, linetype = "22", color = "grey20") +
  facet_wrap(~Bin) +
  scale_fill_viridis_c() +
  labs(x = "Nonrestrictive Solution", y = "Restrictive Solution", 
       title = afe_formula) +
  scale_x_log10(labels = scales::label_log()) + scale_y_log10(labels = scales::label_log()) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    legend.title.position = "top",
    legend.title = element_text(hjust = 0.5)
  )

# For this figure, try reducing bins to Distinguishable and Not Distinguishable

@@

```{r}
rat_physchem <- filter(
  chem.physical_and_invitro.data,
  DTXSID %in% bindf_rat$DTXSID
) |>
  select(
    DTXSID,
    logHenry, logMA, logP, logPwa, logWSol,
    Rat.Clint, Rat.Funbound.plasma,
    # pKa_Accept, pKa_Donor, Chemical.Class,
    MP, MW
  ) |>
  mutate(
    across(starts_with("Rat."), as.numeric)
  )

lombardo_trim <- lombardo |>
  filter(!is.na(DTXSID)) |>
  select(
    DTXSID,
    human_v_dss_l_kg,
    human_cl_m_l_min_kg,
    terminal_t1_2_h,
    hba, hbd,
    moka_ion_state7_4
  ) |>
  mutate(moka_ion_state7_4 = factor(moka_ion_state7_4))

tree_rat <- bindf_rat |>
  left_join(cyp_data, by = join_by(DTXSID == Chemical)) |>
  left_join(rat_physchem, by = join_by(DTXSID)) |>
  # left_join(lombardo_trim, by = join_by(DTXSID)) |>
  select(-c(DTXSID, CAS)) |>
  filter(!Bin %in% "ND_MC_mid") |> # Only one observation, don't want it to throw of tree
  mutate(
    Bin = factor(Bin),
    Bin = forcats::fct_other(
      Bin,
      keep = "Distinguished", 
      other_level = "Not Distinguishable" 
    )
  )

table(tree_rat$Bin) 
# Quite unbalanced, by we can try to upsampling/downsampling

fit_tree_rat <- rpart::rpart(Bin ~ ., data = tree_rat)
rpart.plot::rpart.plot(fit_tree_rat)
preds_tree <- predict(fit_tree_rat, tree_rat, type = "class")
caret::confusionMatrix(preds_tree, tree_rat$Bin, positive = 'Distinguished')

# adjusted Clint   = Clint / Fu_hep... make sure I am using this value
# randomForest < TRY
# Potentially, add uncertainty factor for bins (Rest vs Non)
